---
title: "UFO Annual Compliance Report"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r Load Libraries, echo = F, warning = F, message = F}
shhh <- suppressPackageStartupMessages
shhh(library(tidyverse))
shhh(library(here))
shhh(library(sf))

# source('AIM_report_functions)
set.seed(24)
rm(shhh)

```


```{r Load AIM input data}

files <- paste0(here(), '/data/raw/',
                list.files(paste0(here(), '/data/raw/')))

AIMdata <- read.csv(files[which(str_detect(files, "Query"))] )
Plotdata <- read.csv(files[which(str_detect(files, "Plots"))] )
soildata <- read.csv(files[which(str_detect(files, "SoilHorizon"))] )
PSPSdata <- read.csv(files[which(str_detect(files, "Species.csv"))] )
SPdata <- read.csv(files[which(str_detect(files, "SpeciesRich"))] )
ST <- read.csv(files[which(str_detect(files, "LPIDetail"))])

rm(files)
```


```{r Join the Ecological Site Description data to the AIM data set, echo = F}
ecological_site_d <- Plotdata %>% 
  dplyr::select(Actual.Eco.Site, PrimaryKey)

AIMdata <- left_join(AIMdata, ecological_site_d, by = 'PrimaryKey')
soildata <- left_join(soildata, ecological_site_d, by = 'PrimaryKey')
PSPSdata <- left_join(PSPSdata, ecological_site_d, by = 'PrimaryKey')
SPdata <- left_join(SPdata, ecological_site_d, by = 'PrimaryKey')

# "Plotdata has the actual.eco.site assigned (This was filled in manually)" - TY
```


```{r Annotate Eco site subset ST dataset, echo = F}
STN <-ST %>%
  select(HeightWoody, HeightHerbaceous, SpeciesWoody, SpeciesHerbaceous, PrimaryKey) %>% # Selects the data from the data frame needed
  drop_na(.) %>%  #remove any NA values
  left_join(., ecological_site_d, by = 'PrimaryKey')

rm(ecological_site_d, ST)
# THE PURPOSE OF THIS CHUNK IS CURRENTLY UNLCEAR TO ME, SEEMS SUBSET AND READY FOR ANALYSES
```

```{r Noxious and Nonnoxious Species Comparison by Strata, echo = F}

target <- c("SS","SD","MMS","MC","PJ","GR","SAL","SAN","FORE") 
# Creates a list with strata abbreviations 

#png(filename =
#      paste0(here(), "/results/plts_fg_tab/NoxVsNonnox.png"), 
#    width = 6, height = 4, units = "in", res = 300) 

AIMdata %>%
  select(AH_NoxCover, AH_NonNoxCover, Actual.Eco.Site) %>% 
  filter(Actual.Eco.Site %in% target) %>% 
  pivot_longer(!Actual.Eco.Site, names_to = 'covertype', values_to = 'covervalue') %>% 

  ggplot(., aes(x=covertype, y=covervalue, fill= covertype))+ 
    geom_jitter(alpha=0.2, width = 0.2)+  
    geom_boxplot() +            
    labs(title = "Noxious and Nonnoxious Species Comparison by Strata", 
         x = "Cover Type",
         y = "Average Percent Cover",
         fill = "Cover Type:") +  
    scale_x_discrete(labels= c("Non-noxious", "Noxious"))+   
    scale_fill_manual(values=c("dodgerblue2","firebrick3"),labels = c("Nonnoxious", "Noxious"))+    
    
    theme_classic() +
    facet_wrap(~ Actual.Eco.Site, ncol = 3) +
    theme(axis.text.x = element_text(colour="gray20", angle = 25, hjust=0.75),
          legend.position="bottom", legend.box = "horizontal",
          plot.title = element_text(hjust = 0.5))

# dev.off()  
```

```{r Perennial and Annual Grass}

#png(filename =
#      paste0(here(), "/results/plts_fg_tab/AnualVsPernGrass.PNG"), 
#    width = 6, height = 4, units = "in", res = 300) 

AIMdata %>%
  select(AH_PerenGrassCover, AH_AnnGrassCover, Actual.Eco.Site) %>%
  filter(Actual.Eco.Site %in% target)  %>% 
  pivot_longer(!Actual.Eco.Site, names_to = 'covertype', values_to = 'covervalue') %>% 

  ggplot(., aes(x=covertype,y=covervalue,fill= covertype)) +
    geom_jitter(alpha=0.2, width = 0.2)+
    geom_boxplot() +
    labs(title = "Percent Cover of Annual and Perennial Grass",
         x = "Duration", y="Average Percent Cover",
         colour= "Duration") +
    scale_x_discrete(labels= c("Annual","Perennial")) +
    facet_wrap(.~Actual.Eco.Site, ncol=3) +
    theme_bw() +

    scale_fill_manual(values=c("paleturquoise3", "rosybrown3"),labels = c("Annual", "Perennial")) +
    theme(axis.text.x=element_text(colour="gray20", angle = 25, hjust=0.75), 
          legend.position="bottom", legend.box = "horizontal",
          plot.title = element_text(hjust = 0.5)) 

dev.off()

```

```{r Create Columns of Total Aerial Plant Cover by Functional Group, echo = F}

AIMdata <- AIMdata %>% 
  mutate('Shrub_Cover'   = rowSums(select(., matches("AH.*ShrubCover")), na.rm = T), # subshrub need to ask Ken I have like 12 columns to taylors 4 ???
         'Succulent'     = rowSums(select(., matches("AH.*SucculentCover")), na.rm = T),
         'Tree'          = rowSums(select(., matches("AH.*TreeCover")), na.rm = T),
         'noxiousforb'   = rowSums(select(., matches("AH_Nox.*ForbCover")), na.rm = T),
         'nonnoxiousforb'= rowSums(select(., matches("AH_NonNox.*ForbCover")), na.rm = T)
         ) %>% 
  group_by(Actual.Eco.Site) %>% # calculate the sample size for each stratum
  mutate('StratumSampleSize' = n()) # by group here

```

```{r stratum description, general using TerrADat form "Plots"}
col_order <- c('Ecological Site', 
       "Elevation.min", "Elevation.Q1", "Elevation.mean", "Elevation.Q3", "Elevation.max", "Elevation.sd",
       "AvgPrecip.min", "AvgPrecip.Q1", "AvgPrecip.mean", "AvgPrecip.Q3", "AvgPrecip.max", "AvgPrecip.sd",
           "Slope.min",     "Slope.Q1",     "Slope.mean",     "Slope.Q3",     "Slope.max",     "Slope.sd"
       )

# REED NOTE : THERE IS SOME MASSIVE CLEAN UP WORK TO DO HERE! I THINK SOME PROBLEMS DUE TO CLIPPING ON A RASTER EXTRACT ARE INVOLVED...
# WONDER IF COMPLETE CASES OR OTHER MIGHT HONESTLY BE A GOOD SOLUTION SO THAT THE N() VALUE IS ACCURATE

Plotdata %>% 
  group_by(Actual.Eco.Site) %>%  # Group by ecosite to get stats for individual strata
  mutate('SampleSize' = n()) %>% 
  mutate(across(c(Elevation, AvgPrecip, Slope), # Get stats for these columns 
               list(min = min, mean = mean, max = max, sd = sd), 
                #    Q1 = quantile(.,probs=0.25, na.rm = T), 
                #    Q3 = quantile(.,probs=0.75, na.rm = T)),     
               na.rm = T,
            .names = "{.col}.{.fn}"
  )) %>% 
    mutate(across(c(Elevation, AvgPrecip, Slope), # Get stats for these columns 
               funs(Q1 = quantile(.,probs=0.25, na.rm = T),
                     Q3 = quantile(.,probs=0.75, na.rm = T)), 
               .names = "{.col}.{.fn}"
               )) %>%  # this not working need better cleaning;
  dplyr::select(SampleSize:ncol(.)) %>% 
  distinct(Actual.Eco.Site, .keep_all = T) %>% 
  rename('Ecological Site' = Actual.Eco.Site)  %>% 
  select(col_order) %>% 
  rename_with(~str_remove(., 'Elevation/.|AvgPrecip/.|Slope/.')) %>% 
  knitr::kable(., booktabs = T) %>% 
  kableExtra::add_header_above(c("", "Elevation" = 6, "Annual Precipitation" = 6, 'Slope' = 6))

rm(col_order)
```



```{r Make Summary table of Land form data, echo = F}

Plotdata %>%
  select(Actual.Eco.Site, LandscapeType) %>% 
  mutate(LandscapeType = gsub("[^[:alnum:][:space:]]", "", LandscapeType)) %>% 
  mutate(LandscapeType = snakecase::to_title_case(LandscapeType)) %>% 
  mutate(LandscapeType = case_when(
    LandscapeType == 'Hills Mountains' ~ 'Mountain/Hill',
    LandscapeType == 'Hill Mountain' ~ 'Mountain/Hill', 
    LandscapeType == 'Floodplain Flood Basin' ~ 'Flood plain/Basin',
    TRUE ~ LandscapeType
    )) %>% 
  dplyr::count(Actual.Eco.Site, LandscapeType) %>% 
  rename('Ecological Site' = Actual.Eco.Site) %>% 
  distinct(.keep_all = T) %>% 
  pivot_wider(names_from = LandscapeType, values_from = n, values_fill = 0) %>% 
  select(any_of(c('Mountain/Hill', 'Alluvial Fan', 'Terrace', 'Flood plain/Basin', 'Flat Plain' ))) %>% 
  knitr::kable() 
  
```

# Soils

```{r Make Summary Table of Soils Statisics}

# https://cran.r-project.org/web/packages/soiltexture/vignettes/soiltexture_vignette.pdf

col_orders <- c(
  'Actual.Eco.Site', 'SampleSize', 'Effer', 'Effer Cnt.', 'Texture', 'Texture_Cnt.',
  "RockFragments.min", "RockFragments.Q1", "RockFragments.mean", "RockFragments.Q3", "RockFragments.max", "RockFragments.sd",
  "ESD_PctClay.min", "ESD_PctClay.Q1", "ESD_PctClay.mean", "ESD_PctClay.Q3", "ESD_PctClay.max", "ESD_PctClay.sd"
)

soildata %>% 
  group_by(Actual.Eco.Site) %>% 
  mutate('SampleSize' = n()) %>% 
  select(Actual.Eco.Site, SampleSize, Texture, RockFragments, Effer, ESD_PctClay) %>% 
  add_count(Texture, name = 'Texture_Cnt.') %>% 
  add_count(Effer, name = 'Effer Cnt.') %>% 
  mutate(across(c(RockFragments, ESD_PctClay), # Get stats for these columns 
               list(min = min, mean = mean, max = max, sd = sd), 
            #        Q1 = quantile(., probs=0.25, na.rm = T), 
             #       Q3 = quantile(., probs=0.75, na.rm = T)),     
               na.rm = T,
            .names = "{.col}.{.fn}"
  ))  %>% 
  mutate(across(c(RockFragments, ESD_PctClay), # Get stats for these columns 
               funs(Q1 = quantile(.,probs=0.25, na.rm = T),
                     Q3 = quantile(.,probs=0.75, na.rm = T)), 
               .names = "{.col}.{.fn}"
               )) %>% 
  select(-RockFragments, -ESD_PctClay) %>% 
  distinct(.keep_all = T) %>% 
  dplyr::select(col_orders) %>% 
  rename('Ecological Site' = Actual.Eco.Site) %>% 
  arrange(Texture_Cnt., .by_group = T)# %>% 
  #knitr::kable() #%>% 
  #kableExtra::add_header_above(c("", "Rock Fragment" = 7, "Annual Precipitation" = 13))
  #kableExtra::collapse_rows(columns = 1:4, valign = "top")
  
rm(col_orders)
```

```{r Soil Stability, warning = F, echo = F}

col_orders <- c(
  'Actual.Eco.Site', 'SampleSize', 'Effer', 'Effer Cnt.', 'Texture', 'Texture_Cnt.',
  "RockFragments.min", "RockFragments.Q1", "RockFragments.mean", "RockFragments.Q3", "RockFragments.max", "RockFragments.sd",
  "ESD_PctClay.min", "ESD_PctClay.Q1", "ESD_PctClay.mean", "ESD_PctClay.Q3", "ESD_PctClay.max", "ESD_PctClay.sd"
)

AIMdata %>%
  select(., c(Actual.Eco.Site , matches("SoilStability"))) %>% 
  group_by(Actual.Eco.Site) %>% 
  mutate(across(matches("SoilStability"), # Get stats for these columns 
               list(min = min, median = median, mean = mean, max = max, sd = sd), 
               na.rm = T,
            .names = "{.col}.{.fn}"
  )) %>% # we may want to calculate MODE in here - need defined and troubleshoot
  mutate(across(matches("sd"),
                 ~ .x * 1.282)) %>% 
  rename_with(~gsub('sd', 'Z-score', .x)) %>% # ask ken if he wants 2 SD or why the conversion is carried out
  distinct(Actual.Eco.Site, .keep_all = T) %>%
  dplyr::select(-SoilStability_All, -SoilStability_Protected, -SoilStability_Unprotected) %>% 
  pivot_longer(!Actual.Eco.Site, names_to = 'Variable', values_to = 'Statistic') %>% 
  mutate(Variable = str_remove(Variable, 'SoilStability_')) %>% 
  mutate(Group = str_remove(Variable, "\\.[^.]*$"), .before = Variable) %>% 
  mutate(Variable = str_extract(Variable, "\\.[^.]*$")) %>% 
  pivot_wider(names_from = Variable, values_from = Statistic) %>% 
  rename_with(., ~ str_to_title(gsub(".", " ", .x, fixed = TRUE))) %>% 
  mutate(Group = forcats::fct_relevel(Group, c("Protected", "All", "Unprotected"))) %>% 
  rename(Ecological_Site = 'Actual Eco Site') %>%
  group_by(Ecological_Site) %>%
  arrange(Ecological_Site, Group) %>% 
  knitr::kable(digits = 1) #%>% 
 # kableExtra::collapse_rows(columns = 1, valign = "top")

```

```{r Calculating Surface Cover}

# this is aerial not basal

AIMdata %>%  
  select(BareSoilCover, AH_ForbCover, AH_GrassCover, FH_TotalLitterCover, 
         FH_RockCover, Shrub_Cover, Succulent, Tree, TotalFoliarCover, Actual.Eco.Site) %>% 
  pivot_longer(!Actual.Eco.Site, names_to = 'covertype', values_to = 'covervalue')  %>% 
  mutate(covertype = str_remove_all(covertype, 'AH_|FH_')) %>% 
  mutate(covertype = snakecase::to_title_case(covertype)) %>% 
  
  ggplot(aes(x=covertype,y=covervalue,fill=covertype)) +      
  geom_point(alpha=0.4) +   
  geom_boxplot() +   
  facet_wrap(~Actual.Eco.Site) +
  labs(x="", y="Percent Cover", fill="Cover Type", 
       title = 'Aerial Cover by Stratum')+ 
  theme_classic() +
  theme(axis.text.x = element_blank(), 
        axis.ticks.x=element_blank(),
        plot.title = element_text(hjust = 0.5)
        )  


```

# Vascular Plant Species

#### NOTES 'DOMINANT' IS A RATHER .... TERM, IT SEEMS WE ARE ATTEMPTING TO FUSE SPECIES WHICH ARE BOTH SOMEWHAT UBIQUITUOUS ACROSS ALL SITES WITHIN A STRATUM, AND HAVE A HIGH COVER VALUE. LET'S THINK ABOUT THIS IN THE FUTURE.

```{r Dominant Species by Stratum}

PSPSdata %>%               
  group_by(Actual.Eco.Site) %>% 
  mutate(N_Pts_Stratum = n_distinct(PlotID)) %>% 
  group_by(Actual.Eco.Site, Species) %>%
  add_count(Species, name = 'N_Pts_SPP_Pres') %>%   # Proportion of occurrences of plant by stratum
  mutate(Prop_Pts_spp_pres = N_Pts_SPP_Pres/N_Pts_Stratum) %>% 
  ungroup(Species) %>%
  arrange(desc(Prop_Pts_spp_pres)) %>% 
  slice_head(prop = 0.30) %>% 
  dplyr::select(Actual.Eco.Site, Species, N_Pts_Stratum:Prop_Pts_spp_pres) %>% 
  rename(Ecological_Site = 'Actual.Eco.Site') %>%
  distinct(.keep_all = T) %>% 
  knitr::kable()

```


# ALPHA SPECIES RICHNESS

it looks like we want to have both number of species by stratum with some descriptions of richness (summary ?), and by functional groups
```{r}

PSPSdata %>% 
  group_by(Actual.Eco.Site) %>% 
  mutate(N_Pts = n_distinct(PlotID),
         No_Spp = n()) %>% 
  add_count(GrowthHabitSub, name = 'N_GHS') %>% 
  
  group_by(Actual.Eco.Site, Noxious) %>% 
  add_count(Noxious, name = 'Native_Spp') %>% 
  ungroup() %>% 
  filter(Noxious == 'NO') %>% 
  dplyr::select(Actual.Eco.Site, N_Pts, No_Spp, 
                Native_Spp, GrowthHabitSub, N_GHS) %>%
  distinct() %>% 
  group_by(Actual.Eco.Site) %>% 
  arrange(Actual.Eco.Site, -N_GHS) %>% 
  knitr::kable() # SEE WHAT WE WANT DO TO ABOUT FUNCTIONALLY UNPLACED TAXA
  
PSPSdata %>% 
  group_by(Actual.Eco.Site, PlotID) %>% 
  mutate(Spp_by_Plot = n_distinct(Species)) %>% 
  add_count(Noxious, name = 'Native_Spp') %>% 
  filter(Noxious == 'NO') %>% 
  ungroup() %>% 
  dplyr::select(Actual.Eco.Site, PlotID, Spp_by_Plot, Native_Spp) %>% 
  arrange(-Spp_by_Plot) %>% 
  distinct() %>% 
  ggplot(., aes(x = Spp_by_Plot,fill = Actual.Eco.Site)) +
  geom_density(alpha = 0.8, position = "fill") +
  labs(title = "Distribution of Plant Species Richness at All Plots", 
     x = "Number of Species",
     y = "",
     fill = "Strata:") +  
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

```


### BETA SPECIES RICHNESS

